\documentclass[a4paper,12pt]{report}

\usepackage{amsmath,amsfonts,mathtools}
\usepackage{amssymb}
\usepackage{hyperref}

\begin{document}
\title{PHY293 Abridged}
\author{Aman Bhargava}
\date{September 2019}
\maketitle

\tableofcontents

\section{Introduction}
Prof. Grisouard has put up some pretty solid 
\href{https://github.com/PHY293-Grisouard/Chapters-2019}{notes} written in jupyter notebooks. 
Those offer a great comprehensive guide for the course contents. The goal here is to give
a very concise overview of the things you need to know (NTK) to answer exam questions. Unlike
some of our other courses, you don't need to be very intimately familiar with the derivations
of everything in order to solve the problems (though it certainly doesn't hurt). Think of this 
as a really good cheat sheet.

\chapter{Simple Harmonic Oscillators (SHO's)}
\section{Set Up}
Pretty much the same as covered last year.
$$F_{restorative} = -kx$$
$$F_{net} = ma$$
$$ma + kx = 0$$
We let $\omega_0^2 = k/m$ and rename $a$ to $\ddot{x}$
\section{ODE and Solution}
\paragraph{ODE:}
$$\ddot{x} + \omega_0^2 x = 0$$
\paragraph{Solution:}
$$x(t) = a cos(\omega_0 t + \theta)$$
Solve for 2 unknowns (usually $a$ and $\theta$) based on two initial conditions.

\chapter{Dampened Harmonic Oscillators (DHO's)}
\section{Set Up}
We have all the same forces from last time plus the force of friction.

$$F_{friction} = F_f = -bv$$

\section{ODE and Solutions}
\paragraph{ODE}
We let $\gamma = b/m$ so that:
$$\ddot{x} + \gamma \dot{x} + \omega_0^2 x = 0$$

\subsection{General Solution}

And for underdampened and over dampened, the overall solution looks like:
$$x(t) = ae^{rt}$$
Where $a, r \in \mathbb{C}$, remembering that $$e^{i \theta} = i sin(\theta) + cos(\theta)$$

This is actually really cool, because the existence of a complex component in $r$ is what 
enables oscillation. 

If we populate the initial ODE with the derivatives of our $x(t)$, we get:
$$ar^2e^{rt} + \gamma are^{rt} + \omega_0^2 ae^rt = 0$$
$$r^2 + \gamma r + \omega_0^2 = 0$$

This results in a simple quadratic for which you can solve for $r$ using the quadratic formula.

$$x = \frac{b \pm \sqrt{b^2 - 4ac}}{2a}$$
$$r = \frac{\gamma \pm \sqrt{\gamma^2 - 4w_0^2}}{2}$$

There will be up to 2 $r$ values. Since the solution space is a 2-D vector space, any 
linear combination of vectors (i.e. values of $r$) is still a valid vector.

You might be wondering how $a$ figures into this whole thing. That's just based on initial
conditions, really. 

\subsection{Regieme 1: Underdampened}
If the dampening is weak enough, there will still be \textbf{some} oscillation before the 
oscillator comes to rest. In this case, the general solution is:

$$x(t) = A_0 e^{-\gamma t/2} cos(w_dt + \phi)$$
Where $w_d^2 = w_{dampened}^2 = \omega_0^2 - \frac{\gamma^2}{4}$

You get this by solving for $r$ with the quadratic formula, plugging the two imaginary 
roots in, then taking just the real components because you know that $x(t)$ is real.

\subsubsection{Logarithmic Decrement}
This quantity is the $ln$ of the ratio of 1 peak to the next.

$$\frac{\gamma T_d}{2}$$

Where $T_d$ is the period based on $w_d$.


\subsection{Regieme 2: Heavily Dampened}
In this case both roots are real. This occurs when $\gamma^2-4\omega_0^2 > 0$

Let $r_p, r_m$ be the two real roots. Then the solution is:

$$x(t) = a_pe^{r_pt} + a_me^{r_mt}$$

You can solve for the a values at leisure with two initial conditions.

\subsection{Regieme 3: Critically Dampened}
When the discriminant of the quadratic is zero (i.e. $\gamma^2 = 4\omega_0^2 = $), you
get an overconstrained problem because you have fewer unknowns than you have constraints.
Therefore, the new solution for this exact case is:

$$x(t) = (A + Bt)e^{-\gamma t/2}$$

And it decays the fastest and is pretty cool. Yay.

\section{Energy}
$$E = K + U$$
$$E = \frac{1}{2}mv^2 + \frac{1}{2}kx^2$$
$$\frac{dE}{dt} = -bv^2$$

\subsection{Underdampened Energy}
The equations get really messy so we assume that $\omega_0 >> \gamma/2$ and $\omega_0 = \omega_d$. After 
you sub everything in and use the pythagorean identity to cancel out the squared sine/cosine,
you get:

$$E = \frac{1}{2} kA_0^2e^{-\gamma t}$$

\subsection{Critical and Overdampened Energy}
These ones are really easy, they just decay and that's it. Just remember the equations for 
energy and solve for the velocity equation.

\section{Q-Factor}
$Q$ is a measure of the tendency to oscillate dividied by the tendency to dampen.
$$Q = \frac{\omega_0}{\gamma}$$
Let $\tau = \frac{1}{gamma}$ be proportional to the lifetime of the oscillator. Then 
$n = \frac{\tau}{T_0}$ is the number of cycles in a lifetime.
$$Q = 2\pi n$$


$Q$ is also proportional to the rate at which dampening removes energy from the system. Then
$$E_n = E_0e^{-\gamma t}$$
Represents the amount of mechanical energy in the system at time $t$.
$$\frac{E_n - E_{n+1}}{E_n} = ... = \frac{2\pi}{Q}$$

Therefore $Q$ is the energy in oscillator divided by the amount of energy lost at the
following \textbf{radian}.

\chapter{Forced Oscillators}
\section{Equation of Motion ODE}
$$\ddot{x} + \omega_0^2 x = F_0' cos(\omega t)$$
Where $\omega_0^2 = k/m$, $F_0' = F_0/m$, $\omega$ is the driving frequency.

\section{Solution: Undamped Forced Oscillator}
$$x(t) = A(\omega) cos(\omega t - \delta(\omega))$$
Where
$$\delta(\omega_0) = 0 \,\,if\,\, \omega < \omega_0, else \pi$$
If you shake it too fast, it'll vibrate in exactly opposite phase. 
$$A(w) = |\frac{\omega_0^2}{\omega_0^2 - \omega^2}A_f|$$

\subsection{$\omega = \omega_0$}
Perfect resonance makes amplitude tend toward infinity.
$$x(t) = \frac{A_f \omega_0}{\sqrt{2}} t cos(\omega_0t - \frac{pi}{4})$$

\section{Solution: Forced Damped Oscillators}
Equation of Motion: $\ddot{x} + \gamma \dot{x} + \omega_0^2 x = A_f w_0^2 cos(wt)$

\paragraph{General Solution: } Same as before 
$$x(t) = A(\omega) cos(\omega t - \delta(\omega))$$

\paragraph{Updates to $A$ and $\delta(\omega)$: } 
$$A(\omega) = \frac{\omega_0^2}{\sqrt{\gamma^2\omega^2+(\omega_0^2 - \omega)^2}}A_f$$
$$tan[\delta(\omega)] = \frac{\gamma \omega}{\omega_0^2-\omega^2}$$

\paragraph{Maximum $A(w)$} occurs when $\omega = \omega_0^2 - \frac{\gamma^2}{2}$. In other words, 
$$\omega_{max} = \omega_0^2[1-\frac{1}{2Q^2}]$$

$$\therefore A_{max} = \frac{Q}{\sqrt{1-\frac{1}{4Q^2}}}A_f$$

\section{Transient Response}
When you suddenly start a damped, forced oscillator, you have one mode that is the long-term response and one 
that is the transient response (as if there is no forcing).

$$x_{free} = A_{free} e^{-\gamma t / 2}cos(\omega_d t + \phi)$$
$$x_{forced} = A(\omega) cos(\omega t - \delta(\omega))$$

$x_{free}$ tend toward zero and only exists for the beginning, so it is the transient response. 

\section{Power in FDHO's}
$$\therefore -\dot{E}(w, t) = P(w, t) = bv^2 = bV^2(w)sin^2(wt-\delta)$$
Where $V$ is the amplitude of the velocity.

\subsection{Average Power Per Cycle}
$$\bar{P}(w) = \frac{F_0^2}{2m} \frac{\gamma}{\gamma^2 + (\omega_0^2/\omega - \omega)^2}$$
\paragraph{Maximizing Case: } $\bar{P}$ is maximized when $w = w_0$

\subsection{Power vs. $\omega$}
Let us plot $\bar{P}(w)$ vs. $\omega / \omega_0$. The $\omega_{full width half height} = \omega_{fwhh} = \gamma$

\chapter{Coupled Oscillators}
\paragraph{Main Takeaways}
\begin{itemize}
\item Motion of coupled systems is a linear combination of orthogonal forms of motion.
\item Motion can be found via 2x2 matrix system
\item Each normal mode has a unique frequency common to all elements. 
\item `Eigen' = `normal'.
\item \textbf{Beating} = when frequencies of modes are very close.
\end{itemize}

\subsection{How to Solve Coupled Systems}
From tutorial note: 
\subsubsection{Determine Equations of Motion}
You must give each mass its own coordinate system where deviation in one direction adds to the zero 
coordinate and deviation to the other subtracts from the zero coordinate.

Implement $f = ma$ and $f = kx$ to create equations of motion of each mass depending on its position 
and the position of its neighbouring masses.

Now we have $$m\ddot{x_a} = ...$$ $$...$$ $$m\ddot{x_c} = ...$$

\subsubsection{Assume $\exists$ a normal mode}
This implies that $$x_a = A cos(wt)$$ $$...$$ $$x_c = C cos(wt)$$

All we need to solve for now is $w$, and $A \to C$. To do so, we sub our definitions for $x_a, x_b, and x_c$ 
into our initial EoM, which gives us relationships between $A, B, C, w$ (cosines cancel out!)

Now we just have to put everything in a matrix equation of the form:
$$[...][A, B, C] = [0, 0, 0]$$
$$[M][A, B, C] = [0, 0, 0]$$

which implies that $det(M) = 0$. This will give us $3$ different eigen frequencies (or however many degrees 
of freedom we have). From there, we just have to solve for the coefficients $A, B, C$ via initial conditions, 
remembering that $x_a = Acos(wt)$ and so on!

\subsection{Review: Results from Linear Algebra}
\begin{itemize}
\item $Av = \lambda v$, $(A-\lambda I)v = 0$
\item We get non-trivial solutions for $v$ if $det(A - \lambda I) = 0$
\item A matrix is diagonizable if $\exists n$ eigen values $\lambda$ (sufficient, not necessary)
\item Each solution to the initial $Ax = b$ is unique because the eigen vectors form a basis.
\end{itemize}

\paragraph{Connecting Eigen Things to Real Things}
\begin{itemize}
\item Each $\lambda_i = \omega_i^2$
\end{itemize}

\chapter{Standing Waves}
\paragraph{travelling wave: } a travelling disturbance where no matter travels on average.
\paragraph{Continuum: } infinite coupled oscillators $\to$ infinite modes can arise, characterized 
by nodes and antinodes.

Boundaries give rise to standing waves (think about boundaries on a plucked guitar string).

\section{Expectations and Takeaways}
\begin{enumerate}
\item Wave equation is $$\frac{\partial^2 y}{\partial t^2} = v^2\frac{\partial^2y}{\partial x^2}$$
\item Phase speed equation is $$v = \sqrt{\frac{T}{\mu}}$$
\item \textbf{Seperation of variables} is when you
\item Wavenumbers $k_n$ are basically the number of antinodes in the given mode and angular frequencies $\omega_n$ are frequency at which mode $n$ virbrates, 
wavelengths $\lambda_n$ are the physical length of each wave in the mode $n$.
\item Orthogonality relates how different modes of oscillations in standing waves don't interact.
\item Total energy = sum of energy in each mode of oscillation.
\end{enumerate}

\section{The Taut String}
Properties of a taut string:
\begin{itemize}
\item $tan\theta = \frac{dy}{dx} << 1$ where $x$ is along the string and $y$ is across the string.
\item Vibrations occur in only one plane
\item \textbf{Phase speed } is $v^2 = T/\mu$ with $T$ as \textbf{tension} and $\mu$ as \textbf{mass per unit length of string}.
\end{itemize}

\paragraph{WAVE EQUATION: } $$\frac{\partial^2 y}{\partial t^2} = v^2 \frac{\partial^2y}{\partial x^2}$$

\section{Solution to Wave Equation}
We assume that the solution is of the form $$y(x, t) = f(x)h(t)$$ so that it is seperable and works nicely 
with the wave equation. When we sub this into the wave equation, we get:

Let $$\frac{1}{h(t)} \frac{d^2h(t)}{dt^2} = -w^2 = v^2 \frac{1}{f(x)} \frac{d^^2f(x)}{dx^2}$$

So we can still have a $w^2$ that fits well into our SHO ODE:

$$h(t) = H cos(wt + \phi)$$ $$f(x) = A cos(kx) + B sin(kx)$$ where $k = w/v$

\subsection{Boundary Conditions}
We know that $y(t, 0) = y(t, L) = 0$ because the string is secured on either end. This forces the $A$ term 
in the equation $f(x) = A cos(kx) + B sin(kx)$ to be zero, so now $f(x) = B sin(kx)$. Now $f(kL) = 0$ for 
any natural number $k$. 

$$k_n = \frac{n \pi}{L} \,\,, \,\, n \in \mathbb{N}$$

Recall that $k = w/v$ so now $w_n = \frac{n \pi v}{L}$. 

\subsection{Final Solution}
$$y_n(x, t) = C_n cos(w_n t + \phi_n)sin(k_n x)$$
$$y(x, t) = \sum_{n=1}^{\infty} y_n(x, t)$$

\subsection{Surrounding Math}
We need some tools to do proofs and reasoning about these standing and travelling waves. 
$$f \odot g = \frac{2}{L} \int_0^L f(x)g(x) dx$$
which represents the inner product of the vector space and 
$\delta_{mn}$ which is 1 when $m=n$.
$$sin(k_n x) \odot sin(k_m x) = \delta_{mn}$$

Because they are orthogonal to eachother.

\subsection{Energy Considerations}
$$E_n = \frac{1}{4} \mu L w_n^2 A_n^2$$

\chapter{Travelling Waves}
The real general solution to the wave equation: 
$$\frac{\partial^2 y}{\partial t^2} = v^2 \frac{\partial^2 y}{\partial x^2}$$
$$y(x, t) = f(x-vt) + g(x+vt)$$

It can be shown via trig sums that this is the same as the standing wave solution when you 
put it between two boundaries.

\paragraph{Most common travelling wave function: } $$y(x, t) = A sin(\frac{2 \pi}{\lambda} (x \pm vt))$$
$$= A sin(k(x\pm vt))$$
$$= A sin(kx \pm wt)$$
\begin{itemize}
\item $\lambda$ = wavelength
\item $k$ = wavenumber
\item $\omega^2 = v^2k^2$, k is the wave number.
\item $\frac{2\pi}{w}$ = period
\end{itemize}

\section{Energy Considerations}
$$K = \frac{1}{2} \mu (\frac{\partial y}{\partial t})^2$$
$$U = \frac{1}{2} \mu v^2 (\frac{\partial y}{\partial x})^2$$

$<>$ corresponds to integration over one wavelength, so $<f> = \int_0^{\lambda} f(x) dx$.

To find $<E>$ we can apply the above formulae for $K$ and $U$ to the general solution 
$y(x, t)=A cos(kx-wt)$.

$$<E> = \frac{1}{2} \mu \lambda A^2 w^2$$

\section{Power Considerations}
Since these waves are unbounded travelling entites, it's useful to consider how much 
energy they transport past a certain point per second.

We start with instantaneous energy $E = \mu A^2 w^2 sin^2(kx - wt)$. Then 
$$\bar{P} = \frac{w <E>}{2 \pi} = \frac{\mu \lambda A^2 w^3}{4 \pi} = \frac{\mu A^2 w^2 v}{2}$$

\section{Reflection and Transmission}
When medium changes discontinuously, you get reflection and transmission. The boundary conditions 
at the discontinuity are 
\paragraph{Boundary Condition 1: } matching frequency, amplitude, and phase.
\paragraph{Boundary Condition 2: } transverse (across-string) forces must match.

Therefore we know that 
$$y_T(x, t) = A_T cos(wt - k_t x)$$ $$y_R = A_R cos (wt + k_r x)$$

have the same $wt$ term, and $k_R, k_T > 0$ so they propagate in the right directions. The reflected 
wave is in the same original medium, so its spatial frequency $k_R = k_I = w/v_1$ while $k_T = w/v_2$.

Now the only numbers that aren't known are $A_R, A_T$. According to boundary condition 1, $A_I + A_R = A_T$. 
Now we use boundary condition to solve for 
$$\frac{A_T}{A_I} = \frac{2k_1}{k_1+k_2} = \pmb{\tau_{12}} = \frac{2 \sqrt{\mu_1}}{\sqrt{\mu_1} + \sqrt{\mu_2}} $$ 
$$\frac{A_R}{A_I} = \frac{k_1 - k_2}{k_1 + k_2} = \pmb{\rho_{12}} = \frac{\sqrt{\mu_1} - \sqrt{\mu_2}}{\sqrt{\mu_1} + \sqrt{\mu_2}}$$

\section{travelling Waves in Multiple Dimensions}
The wave equation become: $$\frac{\partial^2 f}{\partial t^2} = v^2\nabla^2f = v^2[\frac{\partial^2 f}{\partial x^2} + \frac{\partial^2 f}{\partial y^2} + \frac{\partial^2 f}{\partial z^2}]$$

\chapter{Dispersion}
For non-dispersive waves (the ones we have been talking about up to this point), the following relationship is true for angular 
frequency, wave number $k$, and velocity $v$: $$\omega^2 = k^2v^2$$

For dispersive waves, you have a non-linear relationship. Therefore different waves disperse at different rates. 

Let's think about the phenomenon of beating: when you have similar frequency waves going past eachother, you get these ``beats'' from 
interference patterns (look up a picture to solidify your understanding!).

Since we have different frequencies going at different rates, the beats now move! Let's have an example: 

\section{Example of Dispersive Waves}
Let the following relationship hold for angular frequency and speed of a wave ($g$ is actually just gravity, this is a weird example)
$$\omega = \sqrt{gk}$$
$$v = \omega/k = \sqrt{g/k}$$

The only thing you can really be asked on the exam is to find the group velocity (according to Shayne, at least): 
$$v_g = \frac{d\omega}{dk} \approx \frac{\Delta \omega}{\Delta k}$$

Group velocity is just the speed of the little wave packets. It's an absolute speed that a stationary observer would see. If you have
two different waves interfereing at $\omega_1$ and $\omega_2$, use the $\Delta$ form to find the speed of the packets!

\chapter{Special Relativity}
\section{Introduction}
An intuitive grasp of this stuff is pretty easy. Just watch the minute physics videos about it, they do a great job.

The key points are that light speed is constant no matter your frame of reference. In other words, it looks like $v_1 + c = v_1$, 
regardless of what $v_1$ is. Therefore, the way you would think to transform a spacetime diagram to be from another person's 
reference frame is not correct. You use this cool Lorentz transformation thing:

\section{Formulae}
\paragraph{The Lorentz factor} comes up all the time: $$\gamma(v) = \frac{1}{\sqrt{1-v^2/c^2}}$$
\subsection{Time Dilation}
Let's say you pass by some dude on a planet at close to the speed of light. Both of you will see the other acting really slowly. In other 
words, \textbf{the moving clock is always observed to be slower than the stationary clock}. Therefore, for this relationship, $\Delta t'$ 
is just \textbf{the other clock}:
$$\Delta t' = \gamma(v)\Delta t$$

\subsection{Length Contraction}
Same thing as above applies with \textbf{the moving object is always observed to be scrunched up more than the stationary object}: 
$$L=\frac{L_0}{\gamma(v)}$$

\subsection{Relativstic Events: Location and Time}
Let's say you're watching a dude running close to the speed of light ($v$) who is about to detonate a cake. If the dude does so at 
position $x$ and time $t$ from his perspective, the position $x_{new}$ and time $t_{new}$ that you observe are given by the following 
equations.
$$t_{new}=\gamma(t-\frac{v^2}{c^2}x)$$
$$x_{new}=\gamma(x-vt)$$

\subsection{Relativistic Energy and Momentum}
$$p=\gamma(v)m_0v$$
$$E = \gamma mc^2$$
$$KE=E-mc^2=\gamma mc^2-mc^2=(\gamma-1)mc^2$$

\paragraph{Invariant Energy: }
$$m_0^2c^4=E-p^2c^2$$

Apparently the doppler shift equation (relativistic) is given, so don't worry too much about that.

\subsection{4-Vectors}
In conventional 3-D space, you can have nice invariant properties of objects, like relative distances. In relatistic 4-D space 
you can have nice invariant 4-vectors that tell you the relative distance of objects and events in 4-D space!

$$\Delta x = (c\Delta t,\Delta x, \Delta y, \Delta z) = (c\Delta t, \pmb{r})$$

\paragraph{The Inner Product} (the thing that is actually invariant) is given by: 
$$A = (A_0, \pmb{A}); \, B = (B_0, \pmb{B})$$
$$A\cdot B=A_0B_0 - \pmb{A}\cdot \pmb{B}$$

Where the bolded A and B are a vector of the three spatial coordinates, the $A_0/B_0$ is the the $c\Delta t$ component of the 4-vector.

The other form of 4-vector is as follows: 

$$...$$



\chapter{Quantum}
\section{Particle Energy, Wave Energy}
Here is a hodge podge of fun properties about waves (and particles, cause they're waves too). Shoutout to De Broglie.
\begin{itemize}
\item Energy of particle/wave with angular frequency: $E = \hslash \omega$
\item Momentum of particle/wave with angular frequency: $p = \hslash \omega/c$. Now is as good a time as any to tell you that $k=\omega/c$, which is really annoying and confusing.
\end{itemize}

\section{Compton Scattering}
When a high energy photon with wavelength $\lambda_0$ collides with an electron, the photon scatters of at angle $\theta$ with respect to its original direction. The wavelength of the photon 
changes accordingly to $\lambda'$

$$\lambda' - \lambda_0 = \frac{R}{mc}(1-\cos\theta)$$
\begin{itemize}
\item $\lambda'$ is the final wavelength of the photon.
\item $\lambda_0$ is the initial wavelength of the photon.
\item $R$ is some constant we're given.
\item $m$ is the mass of the electron.
\item $c$ is the speed of light.
\item $\theta$ is the angle of deflection of the photon with respect to it's initial line of attack.
\end{itemize}

\section{Photoelectric Effect}
If you yeet some photons at a piece of metal, the electrons get ejected if the energy of the light packets is high enough (if the frequency of light is high enough).
$$E_x = \hslash \omega - \phi$$
where $\omega$ is the frequency of light and $\phi$ is the binding energy of the electrons. $E_x$ is the energy of the ejected electrons.

\section{Hydrogen Model}
$$\omega_{mn}=2\pi R_0 c(\frac{1}{n^2} - \frac{1}{m^2})$$

\section{De Broglie Waves}
$$\lambda = \frac{h}{p} = \frac{h}{mv}$$
$$E = \hslash \omega$$
$$p = \hslash k; \,\,\,\,\,\, k = \omega/c$$

\section{Wave-Particle Duality}
\subsection{Uncertainty Principle}
$$\Delta x\Delta p_x \geq 2\pi \hslash$$
$$$$
\subsection{Double Slit Experiment}
If you shine a laser through two slits (or anything really), the two slits will act as sources of the exact same frequency light and they will 
interfere with eachother to create some diffraction patterns on wall. Difference in path length being an integer multiple of the wavelength makes 
constructive interference, difference in path length of half makes for destructive.

$$m \lambda = d\sin \theta$$
Where \begin{itemize}
\item $m$ is some integer (0 is the central node, 1 is the next, etc.)
\item $\lambda$ is the wavelength of light (or particles) coming through.
\item $d$ is the distance between the slits.
\item When the equation holds true for integer value of $m$, you get constructive point. If it's an integer and a half, it's a destructive point.
\end{itemize}


\section{Black Body Radiation}
A body absorbs and emits radiation with the exact same efficiency for any given frequency. A black body is a perfect absorber and perfect emitter. 
$$P=\sigma T^4$$
$P$ is the power per square meter of the emitter surface, $\sigma$ is the emissivity/absorptivity constant, and $T$ is the temperature.

$$f_{max} \propto T$$
In other words, the frequency of radiation that is absorbed/emitted most is proportional to temperature.



\chapter{Matter Waves and The Schrödinger Equation}
This problem is basically just $\pmb{A}\vec{x} = \lambda \vec{x}$ as follows:
$$(\frac{p^2}{2m} + V) \Psi = E\Psi$$
\begin{itemize}
\item $p$ is the momentum
\item $m$ is the mass
\item $V$ is a conservative field of potential energy..
\item $E$ is the energy.
\end{itemize}

\section{Normalization and Probability}
Solve for $N$ such that $$\int_{-\infty}^{\infty} |(N\psi(x))^2| dx = 1$$

The probability of the particle being between two points is just the integral between those of $\int_a^b |\psi(x)^2| dx$ as long as $\psi$ is 
normalized. 

\section{Energy Considerations}
\paragraph{Quantum Kinetic and Potential Energy} are given by: $$U = \frac{\hat{p}^2}{2m}\psi(x)$$
$$K = \frac{\hslash^2}{2m}\nabla^2\psi(x)$$
Where
\begin{itemize}
\item $\hat{p} = -i\hslash \frac{d^2}{dx^2}$
\end{itemize}

\section{Infinite Well}
This is a pretty common question to ask. The high level explanation is as follows:
\begin{itemize}
\item There is a square well of potential energy with infinitely high side walls.
\item The particle is trapped inside. It cannot tunnel out because of the infinitely high walls.
\item At the edges of the well and beyond, the wave function must be 0
\end{itemize}
This is starting to sound a lot like the boundary conditions for standing waves! You can derive this relationship with a bunch
of complicated math, but here's the final formula for the energy of each of the possible standing wave pattern that a particle 
can occupy: 
$$E_n = \frac{n^2\pi^2\hslash^2}{2mL^2}$$
Where $n$ is any positive integer, $m$ is the mass of the particle, $L$ is the length of the well. 

\paragraph{The actual equation} for the wave is as follows: $$\Psi(x) = C\sin(Ax)$$
Where $AL = n\pi$.

\subsection{Superposition}
When you \textbf{measure} the particle, there is only one energy level $n$ that tells you the frequency of the standing wave and, by extension, 
the energy at that level.

Before you observe the particle, it has a superposition of a bunch of $n$ values. However, once you observe the particle, it becomes constrained 
to one $n$ value. This is why the photoelectric effect still has discrete energy levels!

\section{Expectation Values}
Honestly you can get this from our vector calculus course pretty well. We know that the probability distribution is given by $P = |\psi^2|$. From vector calculus, 
we also know how to get the expected values of things from probability distributions. They like to write this with annoying complex conjugates but it's the same thing.
$$\langle x \rangle = \int_{-\infty}^{\infty}\psi^* x \psi dx = \int_{-\infty}^{\infty}|\psi^2| x dx = \int_{-\infty}^{\infty}P(x) x \psi dx$$

So yeah, nothing really to worry about. Same stuff different variable names.

\section{Operators}
Operators are applied to the wave function $\Psi$ to get physical values out of the wave function. For some reason, those values are called eigenvalues. Whatever. 
Here are the operators.
\begin{enumerate}
\item Position operator $\hat{x}$: $\hat{x}\Psi = x\Psi; \hat{r}\Psi=r\Psi$
\item Momentum operator: $\hat{p}$: $\hat{p}_x=-i\hslash\frac{\partial}{\partial x}$
\item Kinetic energy operator: $\hat{K}$: $\hat{K}=-\frac{\hslash^2}{2m}\nabla^2$
\item Total energy operator: $\hat{H}$: $\hat{H} = \hat{K} + \hat{U}$
\item Potential energy is just from any fields of potential energy the particle happens to be riding on.
\end{enumerate}

\section{Schrödinger's For Free Particle}
Remember the general traveling wave equation? Yup, it's back baby!
$$\Psi=Acos(kx-\omega t)+iAsin(kx-\omega t)$$
And here's how you get all those parameter values:
$$p=\hslash k; \,\,\,K=\frac{1}{2}mv^2;\,\,E=\hslash\omega$$
Just remember that $k$ and $K$ are different things ($k$ is the spatial frequency, $K$ is the kinetic energy of the particle). 


\end{document}